---
title: "Group Coupling"
output: html_document
date: "2024-11-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script preprocesses the spindle and slow oscillation data to feed it into the python scripts.
Because these data were analyzed by a team, there are different IDs for the same participant. I united these IDs for the further analysis.

```{r SO read data, echo=FALSE}
base_dir <- "./data/EEG_Events/Events/"

for(i in 1:13){
  if(i==1){
    all_data_SO <- read.csv(paste0("./data/EEG_Events/Events/test_lgipat_test_run1_so_events_datanum_", i,".csv"))
    next
  }
  
  tmp <- read.csv(paste0("./data/EEG_Events/Events/test_lgipat_test_run1_so_events_datanum_", i,".csv"))
  all_data_SO = rbind(all_data_SO, tmp)
}

unique(all_data_SO$channel)

library(tidyr)
library(dplyr)
all_data = all_data_SO
# NegPeak in seconds 
all_data_SO <- all_data_SO %>% select(channel, datasetnum, seconds_begin, seconds_end, seconds_peak_max, seconds_trough_max, duration_seconds, amplitude_peak2trough_max) %>% mutate(NegPeak = seconds_trough_max)

idx_chan <- data.frame(channel = c("F3", "F4", "C3", "C4", "O1", "O2"), idx = c(1:6))
all_data_SO$Annika_Id <- all_data_SO$datasetnum
subjects <- data.frame(Subject_ID = c("13", "12", "11", "10", "9", "8", "7","6","5", "4", "3", "2", "1"), Annika_Id = c("13", "12", "11", "10", "9", "8", "7","6","5", "4", "3", "2", "1"))

all_data_SO$Annika_Id <- as.character(all_data_SO$Annika_Id)

all_data_SO <- merge(all_data_SO, subjects, by = "Annika_Id")
all_data_SO <- merge(all_data_SO, idx_chan, by = "channel")

all_data_SO <- all_data_SO %>% group_by(Subject_ID) %>% arrange(NegPeak,.by_group = T)



```

```{r exclude all doubles, echo=FALSE}
tmp_data = all_data_SO

for(j in 1:6){
for(i in 1:length(tmp_data$NegPeak)){
  if(i >= length(tmp_data$NegPeak)){
    next
  }
  tmp = tmp_data$NegPeak[i+1]-tmp_data$NegPeak[i]
  # check if slow oscilations are less than 1.5 seconds apart
  if (tmp <1.50){ 
    # check which amplitude is bigger and then delete row with the smaller one
    if(tmp_data$amplitude_peak2trough_max[i] < tmp_data$amplitude_peak2trough_max[i+1]){
     tmp_data = tmp_data[-c(i),] 
    }
    else{
      tmp_data = tmp_data[-c(i+1),] 
    }
  }
  
}
}

filt_SO = tmp_data


```


Map the different IDs to each other:
```{r SP read data, echo=FALSE}

base_dir <- "./data/EEG_Events/Events/"

for(i in 1:13){
  if(i==1){
    all_data_SP <- read.csv(paste0("./data/EEG_Events/Events/test_lgipat_test_run1_spindle_events_datanum_", i,".csv"))
    next
  }
  
  tmp <- read.csv(paste0("./data/EEG_Events/Events/test_lgipat_test_run1_spindle_events_datanum_", i,".csv"))
  all_data_SP = rbind(all_data_SP, tmp)
}

unique(all_data_SP$channel)

library(tidyr)
library(dplyr)
all_data = all_data_SP

# NegPeak in seconds 
all_data_SP <- all_data_SP %>% select(channel, datasetnum, seconds_begin, seconds_end, seconds_peak_max, seconds_trough_max, duration_seconds, amplitude_peak2trough_max) %>% mutate(Peak = seconds_peak_max)

idx_chan <- data.frame(channel = c("F3", "F4", "C3", "C4", "O1", "O2"), idx = c(1:6))

all_data_SP$Annika_Id <- all_data_SP$datasetnum
subjects <- data.frame(Subject_ID = c("13", "12", "11", "10", "9", "8", "7","6","5", "4", "3", "2", "1"), Annika_Id = c("13", "12", "11", "10", "9", "8", "7","6","5", "4", "3", "2", "1"))

all_data_SP$Annika_Id <- as.character(all_data_SP$Annika_Id)

all_data_SP <- merge(all_data_SP, subjects, by = "Annika_Id")
all_data_SP <- merge(all_data_SP, idx_chan, by = "channel")

all_data_SP <- all_data_SP %>% group_by(Subject_ID) %>% arrange(Peak,.by_group = T)
```

Exclude all doubling spindles:
```{r SP exclude all doubles, echo=FALSE}
tmp_data = all_data_SP

for(j in 1:6){
for(i in 1:length(tmp_data$Peak)){
  if(i >= length(tmp_data$Peak)){
    next
  }
  tmp = tmp_data$Peak[i+1]-tmp_data$Peak[i]
  if (tmp <0.50){
    # check which amplitude is bigger and then delete row with the smaller one
    if(tmp_data$amplitude_peak2trough_max[i] < tmp_data$amplitude_peak2trough_max[i+1]){
     tmp_data = tmp_data[-c(i),] 
    }
    else{
      tmp_data = tmp_data[-c(i+1),] 
    }
  }
}
}
filt_SP = tmp_data



```

Transform the SO & Spindle events for Python analysis.

```{r transform for python, echo=FALSE}
filt_SO$MidCrossing <- (abs(filt_SO$seconds_trough_max - filt_SO$seconds_peak_max)/2)+filt_SO$seconds_trough_max
# sanity check
# test = filt_SO %>% group_by(datasetnum) %>% mutate(check_SO = ifelse(MidCrossing > seconds_peak_max, 1, 0))
filt_SO$Subject <- filt_SO$Subject_ID
filt_SO$IdxChannel <- filt_SO$idx - 1


write.csv(filt_SO, "./derivatives/Group/slow_waves/SO_events.csv")

filt_SP$IdxChannel <- filt_SP$idx -1
filt_SP$Subject <-filt_SP$Subject_ID
test <- filt_SP %>% group_by(Subject) %>% filter(Subject == 9)
n_occur <- data.frame(table(test$IdxChannel))
n_occur[n_occur$Freq > 1,]
test[test$IdxChannel %in% n_occur$Var1[n_occur$Freq > 1],]

write.csv(filt_SP, "./derivatives/Group/spindles/SP_events.csv")
library(dplyr)
filt_SP %>% group_by(Subject_ID, Annika_Id) %>% summarize(n())

```


## Group coupling analysis
Detect the overlapping events (of slow oscillations and spindles)

```{r coupling, echo=FALSE}

library(dplyr)
sp_raw <- filt_SP
sp_n <- sp_raw %>% group_by(Subject)%>% summarise(count = n())
N <- length(sp_n$Subject)
sw_raw <- filt_SO
sw_n <- sw_raw %>% group_by(Subject)%>% summarise(count = n())
Sub <- sw_n$Subject
sp_n <- sp_n$count
sw_n <- sw_n$count

sp_tmp <- c(sp_raw$Peak)
sw_tmp <- c(sw_raw$NegPeak)


sp <- matrix(, nrow = N, ncol = max(sp_n))
sw <- matrix(, nrow = N, ncol = max(sw_n))

# start values
l = 1
u = sp_n[1]
for(i in 1:N){ # loop through participants
  if(i > 1) {
  u = u+sp_n[i]
  l = l+sp_n[i-1]
  }
  sp[i,1:sp_n[i]] = sp_tmp[l:u]
}

# start values
l = 1
u = sw_n[1]
for(i in 1:N){ # loop through participants
  if(i > 1) {
  u = u+sw_n[i]
  l = l+sw_n[i-1]
  }
  sw[i,1:sw_n[i]] = sw_tmp[l:u]
}

index_coupl <- matrix(, nrow = 0, ncol = 2)
coupl_sp <- sp
# compute coupling
for(sub in 1:N) { # loop over subjects
  for(so in 1:sw_n[sub]) { # iterate over SOs
    curr_so = sw[sub,so]
    for(spindle in 1:sp_n[sub]){# iterate over Spindles
      curr_sp = sp[sub,spindle]
      if ((curr_so+0.75 > curr_sp) & (curr_so-0.75 < curr_sp)){ # check if spindle is within 2 seconds of SO

        index_coupl = rbind(index_coupl, c(Sub[sub], spindle))
        
      } #end of if
    } # end of spindle loop
  }# end of iterate over SOs
} # sub loop ends

sp_n_sum <- sp_n

for(i in 1:length(sp_n)){
  if(i == 1){
    sp_n_sum[i] <- sp_n_sum[i]
  }
  else{
  sp_n_sum[i] <- sp_n_sum[i] + sp_n_sum[i-1]
  }
}

index_coupl <- as.data.frame(index_coupl)

index_coupl$Subject <- index_coupl$V1
index_coupl$index <- index_coupl$V2
index_coupl$coupling <- "coupled"
index_coupl$V2 <- NULL
index_coupl$V1 <- NULL

sp_raw <- sp_raw %>% group_by(Subject) %>% mutate(index = row_number(Subject))

sp_raw <- merge(sp_raw, index_coupl, by = c("Subject", "index"), all = T)

sp_raw$coupling[is.na(sp_raw$coupling)]<- "uncoupled"

write.csv(sp_raw, "./derivatives/Group/spindles/SP_coupling.csv")
```

